---
title: "Analysis of power pose data"
author: "Shravan Vasishth"
date: "1 October 2016"
output:
  github_document: null
  html_document: null
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, comment = "#>", collapse = TRUE, tidy = TRUE)
```

## Load and prepare data

```{r}
## cleaned data
datc<-read.csv("ccy-clean-data.csv",header=TRUE)

# two alternate factor codings for gender
datc$gender_female <- datc$female
datc$gender_male <- relevel(datc$gender_female, "Male")

# testosterone diff
datc$test_diff <- datc$testm2 - datc$testm1

## sanity check: one subject, one row
dim(datc)
length(unique(datc$id))

# drop ineligible and something else as in stata code:
unique(datc$inelig)
unique(datc$anyoutv1)
datc<-subset(datc,inelig!="Ineligible (drop)" & anyoutv1!="Selected")
```

## Examine male and female testosterone levels

```{r}
## subset males and females
males<-subset(datc,female=="Male")
females<-subset(datc,female=="Female")

## initial testosterone:
summary(males$testm1)
summary(females$testm1)

## after treatment:
summary(males$testm2)
summary(females$testm2)
```

## Calculate mean post-treatment testosterone by gender

```{r}
# mean and 95% CI, rounded to 1 digit
mean_se1 <- function(...) round(Hmisc::smean.cl.normal(...), 1)

aggregate(testm2 ~ gender_female + hptreat, datc, mean_se1)

aggregate(testm2 ~ hptreat, datc, mean_se1)
```

By how much did testosterone increase after treatment? Means by gender.

```{r}
## difference scores:
round(with(datc,tapply(testm2-testm1,IND=list(female,hptreat),mean)))

aggregate(test_diff ~ hptreat + gender_female, datc, mean_se1)
```

Means ignoring gender: here is the rough and ready version of fig 3 of the paper.
The effects are a bit smaller in this data-set than the published result, probably because of the statistician's cleaning up of the data.

```{r barplot}
aggregate(test_diff ~ hptreat, datc, mean_se1)

(means<-round(with(datc,tapply(testm2-testm1,IND=hptreat,mean))))
(sds<-with(datc,tapply(testm2-testm1,IND=hptreat,sd)))
n<-length(datc$testm2)
ses<-sds/sqrt(n)

barplot(means,ylim=c(-15,15))
arrows(x0=.75,x1=.75,y0=means[1]-1.96*ses[1],y1=means[1]+1.96*ses[1],angle=90,code=3)
arrows(x0=1.9,x1=1.9,y0=means[2]-1.96*ses[2],y1=means[2]+1.96*ses[2],angle=90,code=3)
```

One could do a two-sample t-test as shown below. Note the high variability in the change in testosterone in low and high pose subjects. There were many cases where testosterone went down post high-power power. 

```{r boxplot}
## calculate difference
diff<-datc$testm2-datc$testm1
## make data frame with differences as DV:
treatment<-datc$hptreat
diff_df<-data.frame(diff=diff,treatment=treatment)
## subset low and high pose subjects' data:
lowdiff<-subset(diff_df,treatment=="Low")
highdiff<-subset(diff_df,treatment=="High")

## graphical summary of distributions:
op<-par(mfrow=c(1,2),pty="s")
boxplot(lowdiff$diff,main="Low",ylim=c(-50,40))
boxplot(highdiff$diff,main="High",ylim=c(-50,40))

## t-test, two sample:
t.test(lowdiff$diff,highdiff$diff)
```

The bummer here is that there is actually no evidence of power posing. 
Cuddy's Ted talk was based on a lie about the data.


## Analyze testosterone using linear models

Now, we will refit the model using various predictors: the initial 
testosterone value, the initial and final cortisone (?) levels, and 
the gender of the subject. The question we ask here is, is post-treatment
testosterone higher for subjects exposed to high vs low power, controlling
for these variables?

```{r}
## center all predictors
datc$ctestm1<-scale(datc$testm1,scale=F)
datc$chptreat<-ifelse(datc$hptreat=="High",1,-1)
datc$cortm1<-scale(datc$cortm1,scale=F)
datc$cortm2<-scale(datc$cortm2,scale=F)
datc$female<-ifelse(datc$female=="Female",1,-1)

## This is the result that Fosse, Cuddy's statistician, found:
summary(m0<-lm(testm2~ctestm1+chptreat+cortm1+cortm2+female,datc))
summary(m0a<-lm(testm2~ctestm1+cortm1+cortm2+female,datc))
## effect on testosterone of power posing:
anova(m0,m0a)

## with interaction with gender, the effect disappears:
summary(m1<-lm(testm2~ctestm1+chptreat+cortm1+cortm2+female+chptreat:female,datc))
summary(m1a<-lm(testm2~ctestm1        +cortm1+cortm2+female+chptreat:female,datc))
anova(m1,m1a)
```

There seems to be no evidence here either that power posing increases testosterone.

## Analyzing the data using change in testosterone

We could also have as dependent measure the change in testosterone in low vs high power subjects. This corresponds to the Fig 3 plot in the paper.

```{r}
datc$change<-datc$testm2-datc$testm1

summary(m2<-lm(change~chptreat+cortm1+cortm2+female,datc))
summary(m2a<-lm(change~cortm1+cortm2+female,datc))
anova(m2,m2a) ## no effect
```

So, the linear model also shows no effect of power posing on change in testosterone.

## Summary

Both the t-test and the likelihood ratio test show that there is *no* evidence in the Cuddy data for power posing raising testosterone. Cuddy's Ted talk was based on a fictional claim.





*** 

Just some other explorations, not originally done by Shravan.

```{r, tidy = FALSE}
# re-express t-test using a linear model
summary(lm(test_diff ~ hptreat, datc))

# look at gender effects
summary(lm(test_diff ~ hptreat * gender_female, datc))
summary(lm(test_diff ~ hptreat * gender_male, datc))
```

```{r line-plots, tidy = FALSE}
# Just doing other explorations
library(ggplot2)
library(dplyr)
library(tidyr)

# Convert to long format
df_long <- datc %>% 
  select(id, gender_male, hptreat, 
         test_1 = testm1, test_2 = testm2, 
         cort_1 = cortm1, cort_2 = cortm2) %>% 
  gather(measure, value, test_1:cort_2) %>% 
  separate(measure, into = c("measure", "time"))

# mean + 95% ci per time per facet
ggplot(df_long %>% filter(measure == "test")) + 
  aes(x = time, y = value) + 
  geom_line(aes(group = id)) + 
  facet_grid(gender_male ~ hptreat) + 
  stat_summary(fun.data = mean_cl_normal, 
               fun.args = list(conf.int = .95), 
               color = "red")

ggplot(df_long %>% filter(measure == "test")) + 
  aes(x = time, y = value) + 
  geom_line(aes(group = id)) + 
  facet_grid(. ~ hptreat) + 
  stat_summary(fun.data = mean_cl_normal, 
               fun.args = list(conf.int = .95), 
               color = "red")

library(lme4)
df_test <- df_long %>% 
  filter(measure == "test") %>% 
  mutate(time_cent = ifelse(time == 1, -.5, .5), 
         time2 = ifelse(time == 1, 0, 1))

# random intercept model
summary(lmer(value ~ time2 * hptreat + (1 | id), df_test))
```

